---
title: "Untitled"
author: "Nishanth Girish"
date: "2023-07-17"
output: word_document
---

########################################################################################################

# ANOVA:
ANOVA is a statistical test that can be used to compare more than two groups.
The analysis of variance is the systematic algebraic procedure of decomposing (i.e.
partitioning) overall variation ( i.e. total variation) in the responses observed in an
experiment into different component of variations such as treatment variation and
error variation. Each component is attributed identifiable cause or source of
variation.

# ASSUMPTIONS:
For the validity of the F-test in ANOVA the following assumptions are made.
1. The effects of different factors (treatments and environmental effects) are
additive in nature.
2. Observations of character under study follow normal distribution
3. Assumption of homogeneity of variance
4. Experimental errors are distributed independently and normally with mean zero
and constant variation i.e. ~ N (0,V)

########################################################################################################


########################################################################################################

A. To check if the given dataset satisfies the assumptions required for ANOVA.

(1) Observations of character under study follow normal distribution:
Testing for Normality: Shapiro Wilks Test:
The Shapiro-Wilk test tests the null hypothesis that the samples come from a
normal distribution against the alternative hypothesis that the samples do not come
from a normal distribution.

```{r}
Circum <- Orange$circumference
shapiro.test(Circum)
```
The conclusion above, is supported by the Shapiro-Wilk test (W = 0.945, p =0.08) which indicates observation are normally distributed.

########################################################################################################


########################################################################################################

(2) Assumption of homogeneity of variance:
Bartlett’s test:
Bartlett’s test tests the null hypothesis that the group variances are equal against
the alternative hypothesis that the group variances are not equal. Bartlett’s test
should be used when the data is normal and Levene’s test should be used when the
data is non-normal.


```{r}
bartlett.test(circumference ~ Tree, data=Orange)
```
Since p values is greater than 0.05, we can conclude that different tree groups are
equal variances.

########################################################################################################


########################################################################################################

(3) Experimental errors are distributed independently and normally with
mean zero and constant variation:
The Shapiro-Wilk test tests the null hypothesis that experimental errors are
distributed normally against the alternative hypothesis that the experimental errors
do not distributed normally.

```{r}
circum.anova.residual <- residuals(aov(circumference ~ Tree, data=Orange))
shapiro.test(circum.anova.residual)
```
Since p values is less than 0.05, we can conclude that experimental errors are not
normally distributed.

```{r}
# Residual Plots:

plot(circum.anova.residual, Orange$Circumference)
```

########################################################################################################

Aim:
Check the equality of average grain yield obtained by varieties A, B and C.

Data set:
Three varieties A, B and C of wheat were compared in the CRD with 6 replications for
each variety. The layout and grain yield in kg/plot are given below. Analyze the data and
state your conclusion.

Procedure:
Hypothesis:
H0: Mean yield of varieties A, B and C are equal

H1: At least two of the varieties have different mean yield

```{r}
data <- read.csv("C:/Users/nisha/Downloads/CRD.csv")
data
```


```{r}
View(data)
str(data)
```
```{r}
varieties <- as.factor(data$Varieties)
names(data)
```
```{r}
boxplot(Yield~Varieties,data=data,xlab= "Varieties",ylab= "Yield")
```
Mean yield of varieties B and C is slightly higher than mean yield of variety A and
observations are normally distributed for variety C as compare to that of Varieties A and
B.

```{r}
model <-aov(Yield~Varieties, data=data)
summary(model)
```
Here since the p-value &gt; 0.05, we fail to reject our H 0. This implies that there is no difference
in mean yield of varieties (ABC).

```{r}
library("report") # Load the package every time you start R

report(model)
```
